Modeling Two-Dimensional Vesicle Hydrodynamics with Hydrophobic Attraction Potential Simulations

### 2021-08-24 doc/main_jfm.tex
Added 4.1, 4.2 (in progress), modified 5.1.2

% Bryan's Marcros
\renewcommand{\aa}{\mathbf{a}}
\newcommand{\bd}{\partial}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\eeta}{\boldsymbol{\eta}}
\newcommand{\FF}{\mathbf{F}}
\newcommand{\GG}{\mathbf{G}}
\newcommand{\JJ}{\mathbf{J}}
\newcommand{\nnu}{\boldsymbol{\nu}}
\newcommand{\ttau}{\boldsymbol{\tau}}
\newcommand{\ssigma}{\boldsymbol{\sigma}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\RR}{\mathbf{R}}
\renewcommand{\SS}{\mathbf{S}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\XX}{\mathbf{X}}
\newcommand{\uu}{\mathbf{u}}
\renewcommand{\vv}{\mathbf{v}}
\newcommand{\yy}{\mathbf{y}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\jump}[1]{[\![ #1 ]\!]}

@@

@@

We have the jump relations for \eqref{eq:uidef}: 
\begin{equation}
\label{eq:DLjump}
\jump{u_i} = \sigma, \quad
\jump{v_i} = 0, \quad
\jump{\nabla u_i} = \frac{\dif \sigma}{\dif s}\ttau,\quad
\jump{\nabla v_i} = \mathbf{0},
\end{equation}
on $\Gamma_i$. We get 
\begin{align*}
  \jump{{\bf T}_2\nnu}   &= \jump{\mathbf{S}(u_i,v_i)\nnu  +\mathbf{S}(v_i,u_i)\nnu} \\
  &= \jump{( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla u_i^T)  \nnu}\\
&= 2\gamma\rho^{-1} \sigma v_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dv_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dv_i }{d\nu} \ttau = \JJ_i,
\end{align*}
Combing  this with \eqref{eq:prejump}  gives  \eqref{eq:recipforcetorque}
as required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Jeffery Orbit}
We use Jeffery orbits to numerically validate our solver for the mobility-problem.  
The background flow is a shear flow $\uu_{\infty}(\xx) = \dot\gamma  \xx \cdot \mathbf{e}_2 \mathbf{e}_1$ with shear rate $\dot\gamma$
and orthogonal unit vectors $\mathbf{e}_1$ and $\mathbf{e}_2$. 
A single elliptical particle is suspended in the fluid with center at the origin.
The Jeffrey orbit  
\begin{equation}
\label{eq:jeff}
\Theta(t) = \tan^{-1}\left(\frac{a}{b}\tan \left(\frac{ab \dot\gamma t}{a^2+b^2}\right)\right)
\end{equation}
gives the angle between the major axis and $\mathbf{e}_2$
where $a$, $b$ are the semi-major and semi-minor axes. 
We are assuming $\Theta(0) = 0$.
Figure~\ref{figure1} shows that the angles coming from the integral equation method \eqref{eq:SKIE}-\eqref{eq:mobility2}
are in complete agreement with the theoretical time course \eqref{eq:jeff}.  Hydrophobic attraction and repulsion are zero for a single particle.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{JefferyOrbit2.eps}
\includegraphics[width=0.4\textwidth]{JefferyOrbit3.eps}
  \caption{The figure compares the theoretical Jeffery orbit for $\Theta(t)$ and the result from the integral equation method. 
    The insets show the simulation setup. The shear rate is $\dot\gamma=0.1$.}
    \label{figure1}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Vesicle in a Shear Flow}

To study the deformation of a Janus-vesicle under a shear flow using the proposed algorithm, 
we adopt a widely tested numerical simulation in both continuum and molecular dynamics models, single vesicle under a shear flow. The well known results indicate that the vesicle has tank-treading behavior when 
there is no viscosity contrast in both fluids inside and outside of the vesicle. Within the framework of the simulation setup, the proposed model provides a feature that the fluid viscosity is constant over the computational domain and it is capable of performing these simulations.

@@

An interesting phenomenon also rises up in the proposed setup where the inter-monolayer slip occurs 
between two leaflets. When a shear rate is large enough, take $\chi = 0.005$ as an example, 
Figure~\ref{figure5} shows the schematic of the test where a particle pair is initially marked and we then track the positions of this particle pair over a period of time. 
In this series of tests, we monitor the mean particle position change and mean angle $\theta(t)$ between particle relative positions with respect to the vesicle center over the time. It is clear to see that the outer leaflet always has larger tangential velocity than the one from the inner leaflet and the mean angle $\theta(t)$ increases. This inter-leaflet sliding phenomenon is consistent with what has been seen
in articles in molecular dynamics simulations.

@@

The inter-monolayer friction coefficient of the Janus-vesicle is calculated from the ratio of the mean friction force and the mean slip velocity. This quantity is given by 
\begin{equation}
b = \frac{\langle F_f\rangle}{\langle\int_{\Sigma_M} R \dot\theta \dif s\rangle},
\end{equation}
%
where $\Sigma_M$ is the curve determined by the midplane and the relative radius $R$ is defined by the distance between the midplane and the centroid of the Janus-vesicle. Hence, the mean rotational velocity 
of the bilayer is given by the integral of $R\dot\theta$ over the midplane curve $\Sigma_M$.
For the friction force $F_f$, we integrate tangential stresses near the bilayer from outer and inner regions then 
extrapolate the force value onto the midplane $\Sigma_M$ to acquire the force difference. 
Table~\ref{table1} provides a series of numerical results for dimensional friction coefficients $b$. As expected,
within the same order, the friction coefficients are independent of the shear rates.

@@

Besides the tests of a Janus-vesicle in a shear flow, an interesting test can also use the proposed model and the setup. Consider a symmetric parabolic flow about the $x$-axis, \cite{Kaoui09} studied shapes of the red blood cell immersed in a parabolic flow using a continuum model. With a similar setup, the background flow for this set of simulations is given by
\begin{equation}
{\bf v}_\infty = v_{max}\left[ 1 - \left( \frac{y}{wR_0}\right)^2 \right]{\bf e}_x,
\end{equation}
%
where $v_{max}$ is the flow strength and $w$ controls the shape of the flow. $R_0$ is defined as the radius of the Janus-vesicle at $t=0$.

>>>>

% Bryan's Marcros
\renewcommand{\aa}{\mathbf{a}}
\newcommand{\bd}{\partial}
\newcommand{\dd}{\mathbf{d}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\eeta}{\boldsymbol{\eta}}
\newcommand{\FF}{\mathbf{F}}
\newcommand{\GG}{\mathbf{G}}
\newcommand{\JJ}{\mathbf{J}}
\newcommand{\nnu}{\boldsymbol{\nu}}
\newcommand{\ttau}{\boldsymbol{\tau}}
\newcommand{\ssigma}{\boldsymbol{\sigma}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\RR}{\mathbf{R}}
\renewcommand{\SS}{\mathbf{S}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\XX}{\mathbf{X}}
\newcommand{\uu}{\mathbf{u}}
\renewcommand{\vv}{\mathbf{v}}
\newcommand{\yy}{\mathbf{y}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\jump}[1]{[\![ #1 ]\!]}

@@

We validated our solver for \eqref{eq:SKIE}-\eqref{eq:mobility2} using a single elliptical particle 
is suspended in a background shear flow; see Section \ref{sec:ves_in_shear}.
Hydrophobic attraction and repulsion are zero for a single particle. 
The angle of the ellipse's major axis coming from the integral equation method was in complete agreement 
with the theoretical, Jeffrey oribit time-course \cite{}.

%The Jeffrey orbit  
%\begin{equation}
%\label{eq:jeff}
%\Theta(t) = \tan^{-1}\left(\frac{a}{b}\tan \left(\frac{ab \dot\gamma t}{a^2+b^2}\right)\right)
%\end{equation}
%gives the angle between the major axis and $\mathbf{e}_2$
%where $a$, $b$ are the semi-major and semi-minor axes. 
%We are assuming $\Theta(0) = 0$.
%

%The background flow is a shear flow $\uu_{\infty}(\xx) = \dot\gamma  \xx \cdot \mathbf{e}_2 \mathbf{e}_1$ with shear rate $\dot\gamma$
%and orthogonal unit vectors $\mathbf{e}_1$ and $\mathbf{e}_2$. 
%A single elliptical particle is suspended in the fluid with center at the origin.
%The Jeffrey orbit  
%\begin{equation}
%\label{eq:jeff}
%\Theta(t) = \tan^{-1}\left(\frac{a}{b}\tan \left(\frac{ab \dot\gamma t}{a^2+b^2}\right)\right)
%\end{equation}
%gives the angle between the major axis and $\mathbf{e}_2$
%where $a$, $b$ are the semi-major and semi-minor axes. 
%We are assuming $\Theta(0) = 0$.
%Figure~\ref{figure1} shows that the angles coming from the integral equation method \eqref{eq:SKIE}-\eqref{eq:mobility2}
%are in complete agreement with the theoretical time course \eqref{eq:jeff}.  Hydrophobic attraction and repulsion are zero for a single particle.

@@

We have the following jump relations for \eqref{eq:uidef}: 
\begin{equation}
\label{eq:DLjump}
\jump{u_i} = \sigma, \quad
\jump{v_i} = 0, \quad
\jump{\nabla u_i} = \frac{\dif \sigma}{\dif s}\ttau,\quad
\jump{\nabla v_i} = \mathbf{0},
\end{equation}
on $\Gamma_i$. We get 
\begin{align*}
  \jump{{\bf T}_2\nnu}   &= \jump{\mathbf{S}(u_i,v_i)\nnu  +\mathbf{S}(v_i,u_i)\nnu} \\
  &= \jump{( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla u_i^T)  \nnu}\\
&= 2\gamma\rho^{-1} \sigma v_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dv_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dv_i }{d\nu} \ttau = \JJ_i,
\end{align*}
Combing  this with \eqref{eq:prejump}  gives  \eqref{eq:recipforcetorque}
as required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Model Parameters}
In the present manuscript, the particles are disks with centre $\aa_i$, and director $\dd_i$.  
The boundary conditions \eqref{eq:SLbc} are $f_i(\xx) = \tfrac{1}{2}(1 + \cos \theta_i)$ where $\theta_i$ is the angle between 
$\xx - \aa_i$ and $\dd_i$.  The particle diameters are $l_0 = 2.5$ nm representing the characteristic thickness of monolayer. 
The other physical parameters are 
\begin{equation}
\label{eq:parameters}
\rho = 4 l_0,\quad \rho_0 = l_0,\quad \gamma = \text{4 pN nm}^{-1},\quad \mu = \text{1 cP}.
\end{equation}
We nondimensionalize the problem through the change of variables 
\begin{equation}
\label{eq:nondimensionalize}
t \mapsto t \text{ ns}, \quad
\xx \mapsto \xx \text{ nm}, \quad 
\xx \mapsto \xx \text{ nm}, \quad 
\uu \mapsto \uu \text{ nm ns}^{-1}, \quad 
  p \mapsto p \text{ pN nm}^2.
\end{equation}
Here $\text{1 cP} = \text{1 pN ns nm}^{-2}$ is for the viscosity of water at room temperature.  
For the rest of the manuscript, the time axes are for ns; the pressure $p$ and stress $\sigma$, $\mathbf{T}^{\mathrm{hydro}}$
are for pN nm$^{-2}$.

\subsection{Permeability}
Vesicle membranes have a large area modulus $K_A$ and changes in surface are give rise to a membrane tension
$\tilde \gamma = K_A(A/A_0 - 1)$ where $A$ is the membrane surface area and $A_0$ is the reference surface area in the non-tensed state. 
Since we are dealing with two-dimensional vesicles, we replace $A$ and $A_0$ with the membrane arc length $L$ and reference length $L_0$
respectively.  Thus, the tension in a stretched cylindrical vesicle membrane is 
\begin{equation}
\label{eq:stretch}
\tilde \gamma = K_A\left(\frac{L}{L_0} - 1 \right).
\end{equation}
A stretched vesicles has a Laplace pressure 
\begin{equation}
\label{eq:LaplacePressure}
\Delta p = \frac{\tilde \gamma}{R} 
\end{equation}
where $\Delta p$ is the difference in internal pressure to the pressure at infinity
and  $R$ is the radius of the circular vesicle.  The factor 
$R^{-1}$ is the total curvature since  the principle curvature in the direction of the cylinder is zero.

The particles in our setup do not abut but rather have small gaps due to repulsive forces.
The gaps allow for some fluid flux across the JP bilayer.  Aqueous flux is quantified by a permeability constant $P$ in the equation
\begin{equation}
\label{eq:perm}
\frac{dA}{dt} = -P  L \Delta p.
\end{equation}
Combining \eqref{eq:stretch}, \eqref{eq:LaplacePressure}, and \eqref{eq:perm}, we can solve the for length $L$ 
\begin{equation}
L_0(L-L(0)) + L_0^2 \ln\left|\frac{L-L_0}{L(0)-L_0}\right| = -4\pi^2 P K_A t
\end{equation}
and compare with our data.

%\subsection{Jeffery Orbit}
%We use Jeffery orbits to numerically validate our solver for the mobility-problem.  
%The background flow is a shear flow $\uu_{\infty}(\xx) = \dot\gamma  \xx \cdot \mathbf{e}_2 \mathbf{e}_1$ with shear rate $\dot\gamma$
%and orthogonal unit vectors $\mathbf{e}_1$ and $\mathbf{e}_2$. 
%A single elliptical particle is suspended in the fluid with center at the origin.
%The Jeffrey orbit  
%\begin{equation}
%\label{eq:jeff}
%\Theta(t) = \tan^{-1}\left(\frac{a}{b}\tan \left(\frac{ab \dot\gamma t}{a^2+b^2}\right)\right)
%\end{equation}
%gives the angle between the major axis and $\mathbf{e}_2$
%where $a$, $b$ are the semi-major and semi-minor axes. 
%We are assuming $\Theta(0) = 0$.
%Figure~\ref{figure1} shows that the angles coming from the integral equation method \eqref{eq:SKIE}-\eqref{eq:mobility2}
%are in complete agreement with the theoretical time course \eqref{eq:jeff}.  Hydrophobic attraction and repulsion are zero for a single particle. 
%
%
%\begin{figure}
%\centering
%\includegraphics[width=0.4\textwidth]{JefferyOrbit2.eps}
%\includegraphics[width=0.4\textwidth]{JefferyOrbit3.eps}
%  \caption{The figure compares the theoretical Jeffery orbit for $\Theta(t)$ and the result from the integral equation method. 
%    The insets show the simulation setup. The shear rate is $\dot\gamma=0.1$.}
%    \label{figure1}
%\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Vesicle in a Shear Flow}
\label{sec:ves_in_shear}
To study the deformation of a Janus-vesicle under a shear flow using the proposed algorithm, 
we adopt a widely tested numerical simulation in both continuum and molecular dynamics models, single vesicle under a shear flow. 
The background flow is a shear flow 
\begin{equation}
\uu_{\infty}(\xx) = \dot\gamma  \xx \cdot \mathbf{e}_2 \mathbf{e}_1
\end{equation}
with shear rate $\dot\gamma$
and orthogonal unit vectors $\mathbf{e}_x$ and $\mathbf{e}_y$ for the horizontal and vertical directions respectively. 
The well known results indicate that the vesicle has tank-treading behavior when 
there is no viscosity contrast in both fluids inside and outside of the vesicle. Within the framework of the simulation setup, the proposed model provides a feature that the fluid viscosity is constant over the computational domain and it is capable of performing these simulations.

@@

\subsubsection{Intermonolayer friction}
An interesting phenomenon also rises up in the proposed setup where the inter-monolayer slip occurs 
between two leaflets. When a shear rate is large enough, take $\chi = 0.005$ as an example, 
Figure~\ref{figure5} shows the schematic of the test where a particle pair is initially marked and we then track the positions of this particle pair over a period of time. 
In this series of tests, we monitor the mean particle position change and mean angle $\theta(t)$ between particle relative positions with respect to the vesicle center over the time. It is clear to see that the outer leaflet always has larger tangential velocity than the one from the inner leaflet and the mean angle $\theta(t)$ increases. This inter-leaflet sliding phenomenon is consistent with what has been seen
in articles in molecular dynamics simulations.

@@

To calculate the inter-monolayer friction coefficient for the apposing JP leaflets, we let  
\begin{equation}
b = \frac{\langle F\rangle}{\langle L \rangle \langle U\rangle },
\end{equation}
%
where $\langle F\rangle = \langle \int_{\mathcal{C}_M} \jump{\ttau \cdot \sigma \cdot \nnu} \,ds\rangle$ 
is the time-averaged, tangential force along the midplane curve $\mathcal{C}_m$. 
The parameters $\langle L\rangle$ and $\langle U\rangle$ are the time-averaged midplane curve length and
intermonolayer slip velocity respectively.

We calculate $\langle F\rangle$ by first evaluating the smooth tangential force along curves a positive distance inside and 
outside $\mathcal{C}_m$ respectively, and then interpolating at zero distance to give a force jump. 
We have made a slight abuse of language: due to the problem being two-dimensional,
the integral $\langle F\rangle$ is actually a force density per length which when divided by length gives the requisite force per area. 
which when divided by length

Table~\ref{table1} provides dimensional friction coefficients $b$ for several shear rates
and the results are in quantitative agreement with 
values previously reported in the literature~\cite{sch-vla-mik2010} and~\cite{denOtter2007}.  
The coefficients generally lie within the range 1 to 2 pN$\cdot$ns nm$^{-2}$ but there are some 
outliers.  These outliers ($\dot\gamma = 0.0025$ ns$^{-1}$ and 
$\dot\gamma = 0.0040$ ns$^{-1}$) correspond to slip velocities that are too low relative to shear rate (see the green and red curve in Figure \ref{figure5}: panel (c)).
The JP vesicle consists of only  58 particles and so it is not surprising that discrete effects produce outliers.

\todo[inline]{If $B$ is the numerically calculated friction coefficient,  then 
$b = B/2.5^3$ pN ns nm$^{-3}$ gives the dimensional value.

If $P$ is the numerically calculated pressure, then $p = P/2.5$ pN nm$^{-2}$ is the dimensional pressure.}

@@

Besides the tests of a Janus-vesicle in a shear flow, an interesting test can also use the proposed model and the setup. Consider a symmetric parabolic flow about the $x$-axis, \cite{Kaoui09} studied shapes of the red blood cell immersed in a parabolic flow using a continuum model. With a similar setup, the background flow for this set of simulations is given by
\begin{equation}
\uu_\infty = v_{max}\left[ 1 - \left( \frac{y}{wR_0}\right)^2 \right]{\bf e}_x,
\end{equation}
%
where $v_{max}$ is the flow strength and $w$ controls the shape of the flow. $R_0$ is defined as the radius of the Janus-vesicle at $t=0$.
